\documentclass{ximera}

% MATH -----------------------------------------------------------
\newcommand{\nats}{\mathbb N}
\newcommand{\ints}{\mathbb Z}
\newcommand{\rats}{\mathbb Q}
\newcommand{\reals}{\mathbb R}
\newcommand{\complex}{\mathbb C}
\newcommand{\powerset}{\mathscr P}

%-------------------------------------------------------------

\pgfplotsset{compat=1.18}
\begin{document}
\begin{abstract}
 \end{abstract}

    \title{Exemplar Examples 5}
    
    \maketitle


Let's consider the first order differential equation $xy^{\,\prime}=x+y-\dfrac{y^2}{x}$ for $x>0$.

\begin{enumerate}

\item Does this equation have any constant solutions?   

\vfill

\item This equation is not linear, not separable, and not Bernoulli.  But it is nonlinear homogeneous.  Let's determine a differential equation $u=\dfrac{y}{x}$ must satisfy.

\vfill
\vfill

% y^{\,\prime}=xu'+u=1+u-u^2 ---> xu'=1-u^2 ----> 1/(1-u^2) u'=1/x



\item Let's solve this equation and then determine a general solution for $y$.

% ln|(1+u)/(1-u)|=ln(x^2)+c ---EXP---> (1+u)/(1-u)=cx^2

% -1+2/(1-u)=cx^2 ---> 2/(1-u)=1+cx^2 ---> u=1 - 2/(1+cx^2).  y=x - 2x/(1+cx^2)


\vfill
\vfill
\vfill

\item  Let's determine the solution to the IVP $xy^{\,\prime}=x+y-\dfrac{y^2}{x}$, $y(1)=0$.

% 0=1-2/(1+c).  c=1.  So y=x - 2x/(1+x^2)

\vspace{1cm}


\end{enumerate}



\newpage


Now let's consider the first order differential equation $y^{\,\prime}=(x+y)^2$.


\begin{enumerate}

\item Does this equation have any constant solutions? 

\vfill

% No!

\item This equation is not linear, not separable, not Bernoulli, and not non-linear homogeneous.  However, let's determine a nice substitution to try in order to transform this differential equation into something nicer.

\vfill
\vfill

% u=x+y so y=u-x and y^{\,\prime}=u'-1=u^2.  So u'=1+u^2 and thus 1/(1+u^2) u'=1.

\item Let's solve that equation, and then determine a general solution for $y$.

% arctan(u)=x+C so u=tan(x+C).  Hence y=tan(x+C)-x is the general solution.

\vfill
\vfill


\item What is the solution to the IVP $y^{\,\prime}=(x+y)^2$, $y(0)=0$?

% y=tan(x)-x

\vfill
\vfill

\item Let's determine the interval $I$ of maximum width on which this solution is valid.

\vspace{1cm}

% (-pi/2,pi/2)


\end{enumerate}



\end{document}