\documentclass{ximera}



% MATH -----------------------------------------------------------
\newcommand{\nats}{\mathbb N}
\newcommand{\ints}{\mathbb Z}
\newcommand{\rats}{\mathbb Q}
\newcommand{\reals}{\mathbb R}
\newcommand{\complex}{\mathbb C}
\newcommand{\powerset}{\mathscr P}

%-------------------------------------------------------------


\begin{document}
\begin{abstract}
 \end{abstract}

    \title{Transformations Part I: Bernoulli}
    
    \maketitle

 Sometimes a substitution is useful to transform a first order non-linear, non-separable differential equation into one we know how to solve.\\

 The first kind we will look at are called $``$Bernoulli equations" and they can be transformed into either a separable de (Method A) or a linear de (Method B).\\

  A non-linear first order de is called a \underline{Bernoulli equation} if it can be written in the form $y'+p(x)y=f(x)y^{r}$ where $r$ is any real number other than $0$ or $1$  (those values of the $r$ make the de linear).\\ \\

 \underline{Method A}:  Find a non-trivial solution $y_1$ to the $``$complementary equation" $y'+p(x)y=0$.   Since you only need ONE solution, pick $y_1=e^{-P(x)}$ where $P(x)$ is ONE antiderivative of $p(x)$.  Then use the substitution $\displaystyle u=\frac{y}{y_1}$ which is equivalent to $y=uy_1$. \\

 $y=uy_1\rightarrow $  $y\,'=u\,y_1\,'+u\,'\,y_1 \rightarrow $  \hspace{1cm}  $(u\,y_1\,'+u\,'\,y_1)+p(x)uy_1=f(x)(uy_1)^r \rightarrow $ \\

  $u(y_1\,' +p(x)y_1)+u\,'\,y_1=f(x) u^r y_1^r\rightarrow $  \hspace{1cm} $u\,'\,y_1=f(x) u^r y_1^r$,  which is separable!\\

 Solve for the general solution for $u$, and then $y=uy_1$. \\ \\ \\


 \underline{Method B}:   One can use the substitution $v=y^{1-r}$ to transform a Bernoulli equation into a linear equation: \\

 $v'=(1-r) y^{-r} y'\rightarrow$  $y'=\frac{1}{1-r} y^{r} v'\rightarrow $  $\frac{1}{1-r} y^{r} v' +p(x)y=f(x)y^{r}\rightarrow $ (dividing both sides by $y^r$) \\

 $\frac{1}{1-r} v' +p(x)y^{1-r}=f(x)\rightarrow$  $\frac{1}{1-r} v'+p(x)v=f(x)$, which is linear!

 Solve for the general solution for $v$, and then $\displaystyle y=v^{\frac{1}{1-r}}$.\\ 


Example:  Let's solve $y'-2xy=(xy)^{3}$, which is a Bernoulli de.  Note:  $y=0$ is a solution to this equation.  Let's look for other solutions.  \\

 I prefer method A:  First find a non-trivial solution $y_1$ to the complementary de $y'-2xy=0$, which you (the viewer) should be able to do!  \\

 One such solution is $y_1=e^{x^2}$.\\

 So the substitution $u=\frac{y}{y_1}$, which is equivalent to $y=uy_1$, in this example is $y=ue^{x^2}$.\\

 $y'=u'e^{x^2}+u\left(e^{x^2}\right)'=u' e^{x^2}+u(2xe^{x^2})\rightarrow $ \\

 $[u' e^{x^2}+u(2xe^{x^2})]-2xue^{x^2}=(xue^{x^2})^3$ by substituting $y$ and $y'$ into the de.\\

 So $u' e^{x^2}=x^3u^3e^{3x^2}$, which is separable! Of course, we have to solve the separable equation.\\

 Separate the variables:  $u^{-3} u' =x^3e^{2x^2}$.\\

 Integrate both sides:  $\displaystyle \int u^{-3}\;du=\int x^3e^{2x^2}\;dx$.  (Integrate by parts on the right side.)\\

 So $\displaystyle -\frac{1}{2u^{2}}=\frac{x^{2}}{4}e^{2x^{2}}-\frac{1}{8}e^{2x^{2}}+C\rightarrow $  \hspace{0.5cm} $\displaystyle \frac{1}{u^{2}}=-\frac{x^{2}}{2}e^{2x^{2}}+\frac{1}{4}e^{2x^{2}}+C$  ($``-2C$" relabeled as $C$.) \\

 By algebra, $\displaystyle \frac{1}{u^{2}}=\frac{(1-2x^2)e^{2x^2}+C}{4}$ ($``4C$ relabeled as $C$).\\

 Then we get $\displaystyle u^2=\frac{4}{(1-2x^2)e^{2x^2}+C}\rightarrow $ \hspace{0.5cm} $\displaystyle u=\pm \frac{2}{\sqrt{(1-2x^2)e^{2x^2}+C}}$.\\

 This (Or $u=0$) is the general solution to the transformed equation.  Finally, using $y=ue^{x^2}$ yields: \\

 $\displaystyle y=\pm \frac{2e^{x^2}}{\sqrt{(1-2x^2)e^{2x^2}+C}}$, OR $y=0$,   which is the general solution to our Bernoulli equation!


\end{document}